---
title: "Project 1"
author: "John Doll"
date: "10/25/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(maps)        
library(ggplot2)
library(mapproj)
library(ggthemes)
library(tidyquant)
library(readxl)
library(ggmap)
library(patchwork)
library(scales)
library(grid)
```

# Map

```{r, warning=FALSE, message=FALSE}
#Ohio map
map.county <- map_data('county')  
ohio.county <- subset(map.county, region=="ohio")

#get corona cases from ohio.gov
ohioDF <- read_csv("https://coronavirus.ohio.gov/static/COVIDSummaryData.csv")
#get ohio population from census.gov source = https://www.census.gov/data/datasets/time-
#series/demo/popest/2010s-counties-total.html#par_textimage_739801612
ohioPop <- read_excel("co-est2019-annres-39.xlsx")

#get the date
Today <- Sys.Date()

#filter out row of totals and mutate variables
ohioDF <- ohioDF %>% 
  filter(Sex != "Total") %>% 
  mutate(OnsetDate = mdy(`Onset Date`)) 
  #filter(Today - OnsetDate <= 30) %>%   ------------------------------------no more filter by time occurence

#Creates dataset grouped by County with case, hospitalization, and death count, used for map and for bottom plots
ohioCountyDF <- ohioDF %>% 
  group_by(County) %>% 
  summarize(ncases = sum(`Case Count`),
            ndead = sum(`Death Due to Illness Count`, na.rm = TRUE),
            nhos = sum(`Hospitalized Count`, na.rm = TRUE)) %>% 
  mutate(CaseCat = cut(ncases,
                       breaks = c(0, 1000, 2500, 5000, 
                                  15000, 30000, 50000))) %>% 
  mutate(HosCat = cut(nhos,
                       breaks = c(0, 100, 500, 
                                  1000, 2000, 2500, 3000))) %>% 
  mutate(DeathCat = cut(ndead,
                       breaks = c(0, 150, 
                                  200, 350, 600, 700))) 


#filter out unneeded rows and columns before combining tables
ohioPop <- ohioPop[c(-1, -2, -3, -4, -93, -94, -95, -96, -97),2]

#combine tables and changing variables
#rename county to region to map it with ohio.county
#https://www.sharpsightlabs.com/blog/rename-columns-in-r/
combinedOhio <- cbind(ohioCountyDF, ohioPop) %>% 
  mutate(countyPop = as.double(`...2`)) %>% 
  select(!`...2`) %>% 
  mutate(rate = (ncases/countyPop) * 1000) %>% 
  mutate(County = str_to_lower(County)) %>% 
  rename(subregion = County)

#combine all data sets into one for state map
rateDB <- merge(ohio.county, combinedOhio, by.x = "subregion", by.y = "subregion")
  
#ohio map with continuous shading by total cases //Need to be categorical?
ohioMap <- ggplot(data = rateDB, aes(x=long,y=lat,group = group, fill=CaseCat)) +
  geom_polygon(color = "black") +
  scale_fill_brewer() +
  guides(fill = FALSE) +
  theme_map()
```

# Bottom panels

```{r, warning=FALSE, message=FALSE}
#Bottom left case graph which displays the top 20 counties by cases
Cases <- ggplot(data = ohioCountyDF[tail(order(ohioCountyDF$ncases), 20), ], aes(y=reorder(County, ncases), x=ncases, fill = CaseCat)) +
  geom_col() +
  scale_fill_brewer() +
  guides(fill = FALSE) +
  geom_text(aes(label = ncases), hjust = -.2) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(hjust = 0.5), panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlim(0, max(ohioCountyDF$ncases + 11000)) +
  labs(y = "", x = "", title = "Case Count") 

#Bottom middle hospitalization graph which displays the top 20 counties by cases
Hosp <- ggplot(data = ohioCountyDF[tail(order(ohioCountyDF$ncases), 20), ], aes(y=reorder(County, ncases), x=nhos, fill = HosCat)) +
  geom_col() +
  scale_fill_brewer() +
  guides(fill = FALSE) +
  geom_text(aes(label = nhos), hjust = -.2) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), plot.title = element_text(hjust = 0.5), panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlim(0, max(ohioCountyDF$nhos + 800)) +
  labs(y = "", x = "", title = "Hospitalization Count") 

#Bottom right death graph which displays the top 20 counties by cases
Deaths <- ggplot(data = ohioCountyDF[tail(order(ohioCountyDF$ncases), 20), ], aes(y=reorder(County, ncases), x=ndead, fill = DeathCat)) +
  geom_col() +
  scale_fill_brewer() +
  guides(fill = FALSE) +
  geom_text(aes(label = ndead), hjust = -.2) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), plot.title = element_text(hjust = 0.5), panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlim(0, max(ohioCountyDF$ndead + 550)) +
  labs(y = "", x = "", title = "Death Count") 

#combine all three graphs into one object
bottom <- Cases + Hosp + Deaths
```

# Side Panels

```{r, warning = FALSE, message = FALSE}
#Dataset for right panel of four charts
ohioCasesDF <- ohioDF %>% 
  group_by(OnsetDate) %>% 
  summarize(ncases = sum(`Case Count`),
            ndead = sum(`Death Due to Illness Count`, na.rm = TRUE),
            nhos = sum(`Hospitalized Count`, na.rm = TRUE))

#breaks for dates on x-axis graph
datebreaks <- seq(as.Date("2020-01-01"), as.Date("2020-11-01"), by = "2 months")

#right top case graph
casesBar <- ggplot(data = ohioCasesDF, aes(x = OnsetDate, y = ncases)) +
  geom_col(color = "blue") +
  scale_x_date(breaks = datebreaks, labels = date_format("%b %d")) +
  scale_fill_brewer() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5), panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.subtitle = element_text(hjust = 0.5, size = 20, color = "red")) +
  labs(y = "", x = "", title = "Cases", subtitle = paste(sum(ohioCasesDF$ncases))) 

#right upper middle hospitalization graph
hosBar <- ggplot(data = ohioCasesDF, aes(x = OnsetDate, y = nhos)) +
  geom_col(color = "blue") +
  scale_x_date(breaks = datebreaks, labels = date_format("%b %d")) +
  scale_fill_brewer() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5), panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.subtitle = element_text(hjust = 0.5, size = 20, color = "red")) +
  labs(y = "", x = "", title = "Hospitalizations", subtitle = paste(sum(ohioCasesDF$nhos))) 

#right lower middle death graph
deadBar <- ggplot(data = ohioCasesDF, aes(x = OnsetDate, y = ndead)) +
  geom_col(color = "blue") +
  scale_x_date(breaks = datebreaks, labels = date_format("%b %d")) +
  scale_fill_brewer() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5), panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.subtitle = element_text(hjust = 0.5, size = 20, color = "red")) +
  labs(y = "", x = "", title = "Deaths", subtitle = paste(sum(ohioCasesDF$ndead))) 

Today <- Sys.Date()

ohioRec <- ohioCasesDF %>% 
  filter(OnsetDate < Today - 10) %>% 
  mutate(recovered = ncases - ndead) 

#right bottom graph for presumed recoveries
recBar <- ggplot(data = ohioRec, aes(x = OnsetDate + 10, y = recovered)) +
  geom_col(color = "blue") +
  scale_x_date(breaks = datebreaks, labels = date_format("%b %d")) +
  scale_fill_brewer() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5), panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.subtitle = element_text(hjust = 0.5, size = 20, color = "red")) +
  labs(y = "", x = "", title = "Presumed Recovered", subtitle = paste(sum(ohioRec$recovered)))

right <- casesBar / hosBar / deadBar / recBar

```

# Pulling it all together

```{r, warning=FALSE, message=FALSE}
pushViewport(viewport(layout = grid.layout(9,3))) #, heights = unit(1, "inches"), width = unit(1, "inches"))))
print(ohioMap, vp = viewport(layout.pos.row = 1:4,
                             layout.pos.col = 1:2))
print(casesBar, vp = viewport(layout.pos.row = 1,
                              layout.pos.col = 3))
print(hosBar, vp = viewport(layout.pos.row = 2,
                              layout.pos.col = 3))
print(deadBar, vp = viewport(layout.pos.row = 3,
                              layout.pos.col = 3))
print(recBar, vp = viewport(layout.pos.row = 4,
                              layout.pos.col = 3))
#print(right, vp = viewport(layout.pos.row = 1:4,
 #                          layout.pos.col = 3))
print(Cases, vp = viewport(layout.pos.row = 5:9,
                              layout.pos.col = 1))
print(Hosp, vp = viewport(layout.pos.row = 5:9,
                              layout.pos.col = 2))
print(Deaths, vp = viewport(layout.pos.row = 5:9,
                              layout.pos.col = 3))


```


Deliverables:

-dashboard
-what did you do: the data frames that I built were these, i downlaoded the ohio data from here, then i filtered out these columns, have someone be able to reproduce what you did, cite references

-citations of libraries
-other resources cited includes website links
-dashboard, processes, citations, then code (whatever order is mentioned in assignment)

-save object as pdf, then import it in later
-used courier for code when copying it to word




